### 2024-04-02 paper/paper.qmd
Update

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Qin He
thanks: "Code and data are available at: https://github.com/qinheinfo/Canada_inflation/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-02 scripts/00-simulate_data.R

@@ -1,16 +1,18 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Install the dependencies required for this project
+# Author: Qin He
+# Date: April 1, 2024
+# Contact: thisisheqins@gmail.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: You must also have the R language installed
+# GitHub: https://github.com/qinheinfo/Canada_inflation/tree/main
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(tibble)
+library(readr)
+library(here)
 
 #### Simulate data ####
 # [...ADD CODE HERE...]

### 2024-04-03 outputs/paper.qmd
Test

--- 2024-04-02 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Analyzing Inflation Dynamics in Canada During and After the COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from 2019 to 2023"
+format: pdf
 author: 
 - Qin He
-thanks: "Code and data are available at: https://github.com/qinheinfo/Canada_inflation/tree/main"
+abstract: "The COVID-19 pandemic has left a significant imprint on global economic structures, influencing inflation rates worldwide.
+Canada, like many countries, experienced fluctuating inflation rates during and post-pandemic, marking a period of economic instability and adjustment.
+This paper delves into the inflation dynamics in Canada from 2019 to 2023, with a special focus on the consumer price index (CPI) as a reflection of the cost of goods and services impacting the everyday lives of Canadians.
+It explores the trajectory of the overall inflation rate, which peaked in 2023, highlighting the role of various sectors, including food, housing, and energy, in driving these trends.
+By analyzing data from Statistics Canada, this study provides insights into how the pandemic and subsequent recovery phases have affected inflation.
+It examines the interplay between supply chain disruptions, monetary policy responses, and changes in consumer behavior, culminating in a comprehensive overview of inflation's impact on the Canadian economy.
+This analysis not only sheds light on the inflationary trends but also contributes to understanding the broader economic implications of the COVID-19 pandemic on Canada."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
-number-sections: true
+editor: visual
+toc: TRUE
+toc-depth: 1
+number-sections: TRUE
 bibliography: references.bib
+thanks: "See the GitHub repository supporting this analysis: https://github.com/qinheinfo/Canada_inflation"
 ---
@@
 #| message: false
+#| echo: false
+#| output: false
 
@@
 library(palmerpenguins)
+library(lubridate)
+library(dplyr)
+library(knitr)
+library(janitor)
+library(scales)
+library(RColorBrewer)
+library(ggplot2)
+library(kableExtra)
+library(here)
+library(arrow)
+
+# Fetch the the clean inflation data
+inflation_data <- arrow::read_parquet(here::here("----"), show_col_types = FALSE)
+
+# Fetch clean CPI grocery item data
+grocery_data <- arrow::read_parquet(here::here("----"), show_col_types = FALSE)
+```
+
 ```
@@
 
-# Model
+# Data
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Measurement
+xxx
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+## Data Analysis
 

### 2024-04-03 scripts/00-simulate_data.R

@@ -9,13 +9,49 @@
 
 
 #### Workspace setup ####
-library(tidyverse)
 library(tibble)
 library(readr)
 library(here)
 
-#### Simulate data ####
-# [...ADD CODE HERE...]
+
+#### Function to simulate data ####
+simulate_cpi_data <- function(years, size, start_value, growth_rate) {
+  set.seed(853) # Ensure reproducibility
+  year_sequence <- seq(from = min(years), to = max(years))
+  values <- numeric(length(year_sequence))
+  
+  values[1] <- start_value
+  for (i in 2:length(year_sequence)) {
+    # Simulate yearly growth based on provided growth rate
+    growth <- rnorm(1, mean = growth_rate, sd = growth_rate / 10)
+    values[i] <- values[i - 1] * (1 + growth)
+  }
+  
+  data <- tibble(
+    REF_DATE = rep(year_sequence, each = size / length(year_sequence)),
+    GEO = "Canada",
+    DGUID = "2016A000011124",
+    `Products and product groups` = "All-items",
+    UOM = "2002=100",
+    UOM_ID = 17,
+    SCALAR_FACTOR = "units",
+    SCALAR_ID = 0,
+    VECTOR = "v41693271",
+    COORDINATE = 2.2,
+    VALUE = rep(values, each = size / length(year_sequence)),
+    STATUS = NA,
+    SYMBOL = NA,
+    TERMINATED = NA,
+    DECIMALS = 1
+  )
+  
+  return(data)
+}
 
 
+#### Simulate Data for 2019-2023 ####
+simulated_cpi_data <- simulate_cpi_data(years = c(2019, 2020, 2021, 2022, 2023), size = 20, start_value = 136, growth_rate = 0.03)
 
+
+#### Save Simulated Data ####
+write_csv(simulated_cpi_data, here::here("inputs/simulated_cpi_data.csv")) # Save to inputs/data

### 2024-04-03 scripts/02-data_cleaning.R

@@ -1,44 +1,46 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Install the dependencies required for this project
+# Author: Qin He
+# Date: April 1, 2024
+# Contact: thisisheqins@gmail.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: You must also have the R language installed
+# GitHub: https://github.com/qinheinfo/Canada_inflation/tree/main
 
-#### Workspace setup ####
+
+#### Workspace set-up ####
 library(tidyverse)
+library(dplyr)
+library(here)
+library(readr)
+library(janitor)
+library(arrow)
+library(lubridate)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
-
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+#### Import and filter the raw data ####
+raw_data <- read_csv(here::here("data/raw_data/Raw data.csv"), show_col_types = FALSE) %>%
+  clean_names() %>%
+  select(ref_date, products_and_product_groups, value) %>%
+  # Assuming 'All-items' is a category, and adding placeholders for specific products
+  filter(products_and_product_groups %in% c("All-items", "PLACEHOLDER_FOR_SPECIFIC_PRODUCTS")) %>%
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
[…]

### 2024-04-03 outputs/paper.qmd
test

@@
 The remainder of this paper is structured as follows. @sec-data....
-
-# Data {#sec-data}
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
 

### 2024-04-03 scripts/02-data_cleaning.R

@@ -21,26 +21,18 @@ library(lubridate)
 raw_data <- read_csv(here::here("data/raw_data/Raw data.csv"), show_col_types = FALSE) %>%
   clean_names() %>%
   select(ref_date, products_and_product_groups, value) %>%
-  # Assuming 'All-items' is a category, and adding placeholders for specific products
-  filter(products_and_product_groups %in% c("All-items", "PLACEHOLDER_FOR_SPECIFIC_PRODUCTS")) %>%
   mutate(
     value = as.numeric(value),
-    year = as.integer(ref_date)
+    year = year(ymd(ref_date)) # Convert ref_date to year
   ) %>%
   drop_na(value) # Drop rows where 'value' is NA
 
-# Example placeholder for specific products could be replaced with actual product names like "Milk", "Bread", etc.
 
-#### Optional: Aggregate data ####
-# If your data has multiple entries per product group per year, you may want to aggregate it.
-# This step averages the 'value' by 'year' and 'products_and_product_groups', adjust as necessary.
-aggregated_data <- raw_data %>%
-  group_by(year, products_and_product_groups) %>%
-  summarise(value = mean(value, na.rm = TRUE), .groups = 'drop')
+# Ensure this is the correct variable containing cleaned or aggregated data
+cleaned_data <- raw_data # Placeholder: replace with actual cleaned/aggregated data variable if different
+
 
 #### Save cleaned data ####
 # Adjust the file path and format as needed
 write_csv(aggregated_data, here::here("data/cleaned_data.csv"))
-# For Parquet format
-write_parquet(aggregated_data, here::here("data/cleaned_data.parquet"))
 

### 2024-04-04 scripts/02-data_cleaning.R
data

@@ -21,12 +21,7 @@ library(lubridate)
 raw_data <- read_csv(here::here("data/raw_data/Raw data.csv"), show_col_types = FALSE) %>%
   clean_names() %>%
   select(ref_date, products_and_product_groups, value) %>%
-  mutate(
-    value = as.numeric(value),
-    year = year(ymd(ref_date)) # Convert ref_date to year
-  ) %>%
   drop_na(value) # Drop rows where 'value' is NA
-
 
 # Ensure this is the correct variable containing cleaned or aggregated data
 cleaned_data <- raw_data # Placeholder: replace with actual cleaned/aggregated data variable if different
@@ -34,5 +29,5 @@ cleaned_data <- raw_data # Placeholder: replace with actual cleaned/aggregated d
 
 #### Save cleaned data ####
 # Adjust the file path and format as needed
-write_csv(aggregated_data, here::here("data/cleaned_data.csv"))
+write_csv(cleaned_data, here::here("data/cleaned_data.csv"))
 

### 2024-04-04 outputs/paper.qmd
Update

@@
 title: "Analyzing Inflation Dynamics in Canada During and After the COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from 2019 to 2023"
-format: pdf
+format: html
 author: 
@@
 
-# Fetch the the clean inflation data
-inflation_data <- arrow::read_parquet(here::here("----"), show_col_types = FALSE)
-
-# Fetch clean CPI grocery item data
-grocery_data <- arrow::read_parquet(here::here("----"), show_col_types = FALSE)
-```
-
+# Fetch clean CPI product item data
+product_data <- read.csv(here::here("data/cleaned_data.csv"), stringsAsFactors = FALSE)
 ```
@@
 
-Our results are summarized in @tbl-modelresults.
+## Food
 
 ```{r}
+#| messages: false
 #| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Food" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Food', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    title = "Trend of Value of Food and All-items from 2019 to 2023",
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-2
+#| fig-cap: "Comparison of Food Value and All-items Value Over Time"
+#| fig-align: center
+#| fig-width: 10
+#| fig-height: 3
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Food', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, Food, `All-items`) |> 
+  rename(
+    `Food Value` = Food, 
+    `All-items Value` = `All-items`
   )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+wide_data_display |>
+  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" ) |>
+  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
 ```
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-04-04 outputs/paper.tex
Update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Analyzing Inflation Dynamics in Canada During and After the COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from 2019 to 2023},
  pdfauthor={Qin He},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Analyzing Inflation Dynamics in Canada During and After the
COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from
2019 to 2023\thanks{See the GitHub repository supporting this analysis:
https://github.com/qinheinfo/Canada\_inflation}}
\author{Qin He}
\date{April 4, 2024}

\begin{document}
\maketitle
\begin{abstract}
The COVID-19 pandemic has left a significant imprint on global economic
structures, influencing inflation rates worldwide. Canada, like many
countries, experienced fluctuating inflation rates during and
post-pandemic, marking a period of economic instability and adjustment.
This paper delves into the inflation dynamics in Canada from 2019 to
2023, with a special focus on the consumer price index (CPI) as a
reflection of the cost of goods and services impacting the everyday
lives of Canadians. It explores the trajectory of the overall inflation
rate, which peaked in 2023, highlighting the role of various sectors,
including food, housing, and energy, in driving these trends. By
analyzing data from Statistics Canada, this study provides insights into
how the pandemic and subsequent recovery phases have affected inflation.
It examines the interplay between supply chain disruptions, monetary
policy responses, and changes in consumer behavior, culminating in a
comprehensive overview of inflation's impact on the Canadian economy.
This analysis not only sheds light on the inflationary trends but also
contributes to understanding the broader economic implications of the
COVID-19 pandemic on Canada.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{1}
\tableofcontents
}
\section{Introduction}\label{introduction}

You can and should cross-reference sections and sub-sections. We use R
Core Team (2023) and Wickham et al. (2019).

The remainder of this paper is structured as follows.
\textbf{?@sec-data}\ldots.

\section{Data}\label{data}

\subsection{Measurement}\label{measurement}

xxx

\subsection{Data Analysis}\label{data-analysis}

\section{Results}\label{results}

\subsection{Food}\label{food}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-2-1.pdf}

\begin{figure}

\centering{

\begingroup\fontsize{10}{12}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Year & Food Value & All-items Value\\
\midrule
2019 & 150.2 & 136.0\\
2020 & 153.7 & 137.0\\
2021 & 157.5 & 141.6\\
2022 & 171.5 & 151.2\\
2023 & 184.3 & 157.1\\
\bottomrule
\end{longtabu}
\endgroup{}

}

\caption{\label{fig-2}Comparison of Food Value and All-items Value Over
Time}

\end{figure}%

\section{Discussion}\label{discussion}

\subsection{First discussion point}\label{sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

\subsection{Second discussion point}\label{second-discussion-point}

\subsection{Third discussion point}\label{third-discussion-point}

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Weaknesses and next steps should also be included.

\newpage

\appendix

\section*{Appendix}\label{appendix}
\addcontentsline{toc}{section}{Appendix}

\section{Additional data details}\label{additional-data-details}

\section{Model details}\label{sec-model-details}

\subsection{Posterior predictive
check}\label{posterior-predictive-check}

In \textbf{?@fig-ppcheckandposteriorvsprior-1} we implement a posterior
predictive check. This shows\ldots{}

In \textbf{?@fig-ppcheckandposteriorvsprior-2} we compare the posterior
with the prior. This shows\ldots{}

\subsection{Diagnostics}\label{diagnostics}

\textbf{?@fig-stanareyouokay-1} is a trace plot. It shows\ldots{} This
suggests\ldots{}

\textbf{?@fig-stanareyouokay-2} is a Rhat plot. It shows\ldots{} This
suggests\ldots{}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-rohan}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\end{CSLReferences}

\end{document}

### 2024-04-18 outputs/paper.qmd
update

@@
 title: "Analyzing Inflation Dynamics in Canada During and After the COVID-19 Pandemic: A Deep Dive into Consumer Price Index Movements from 2019 to 2023"
-format: html
+format: pdf
 author: 
@@
 ## Measurement
-xxx
 
 ## Data Analysis
+This paper uses data from both Statistics Canada’s Consumer Price Index (CPI) and the Bank of Canada’s Inflation Rate measurements.
+Within the CPI, the paper specifically examines 6 variables representing different food items: bacon (per 500 grams), eggs (per 1 dozen), chicken breasts (per kilogram), milk (per 2 liters), frozen peas (per 750 grams), and pasta sauce (per 650 milliliters).
+These 6 food items were 3selected as a representation of various food types and groups, allowing an analysis of meats, dairy, frozen goods, and shelf-stable food.
+The CPI displays a price of each food item by month of a given year, and was subsequently cleaned to find the average of a given year.
+Limitations in the history of the data only allow for analysis from 2017 up to 2022.
+Once the data was cleaned, a cost rate (the difference in price of a given year against the previous year) was calculated to show the change in price in order to measure the food item against inflation.
+It is important to note that, since the data only extends to 2017, a cost rate analysis from the years 2018 to 2022.
+The rate of inflation measurements similarly recorded data month-to-month and required cleaning to show the yearly average.
+Although the inflation rate data set provided a far more robust history, only the years 2017 to 2022 were included in this analysis as inflation was only used as a comparison against food cost rates.
 
@@
 #| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Food Value and All-items Value Over Time"
 
@@
   labs(
-    title = "Trend of Value of Food and All-items from 2019 to 2023",
     y = "Value",
@@
 #| warning: false
-#| label: fig-2
-#| fig-cap: "Comparison of Food Value and All-items Value Over Time"
-#| fig-align: center
-#| fig-width: 10
-#| fig-height: 3
+#| label: tbl-1
+#| tbl-cap: "Comparison of Food Value and All-items Value Over Time"
 # Read and prepare the dataset for displaying in a table format
@@
 # Render the table using knitr::kable and kableExtra::kable_styling
-wide_data_display |>
-  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" ) |>
-  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
+kable(wide_data_display)
+```
+
+## Household operations, furnishings and equipment
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Household Value and All-items Value Over Time"
+
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Food" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Household operations, furnishings and equipment', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: tbl-1
+#| tbl-cap: "Comparison of Household Value and All-items Value Over Time"
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Household operations, furnishings and equipment', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, 'Household operations, furnishings and equipment', `All-items`) |> 
+  rename(
+    `Household Value` = 'Household operations, furnishings and equipment', 
+    `All-items Value` = `All-items`
+  )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+kable(wide_data_display)
+```
+
+## Clothing and footwear
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Clothing Value and All-items Value Over Time"
+
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Clothing and footwear" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Clothing and footwear', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: tbl-1
+#| tbl-cap: "Comparison of Clothing Value and All-items Value Over Time"
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Clothing and footwear', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, 'Clothing and footwear', `All-items`) |> 
+  rename(
+    `Clothing Value` = 'Clothing and footwear', 
+    `All-items Value` = 'All-items'
+  )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+kable(wide_data_display)
+```
+
+## Gasoline
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Gasoline Value and All-items Value Over Time"
+
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Gasoline" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Gasoline', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: tbl-1
+#| tbl-cap: "Comparison of Gasoline Value and All-items Value Over Time"
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Gasoline', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, 'Gasoline', `All-items`) |> 
+  rename(
+    `Gasoline Value` = 'Gasoline', 
+    `All-items Value` = 'All-items'
+  )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+kable(wide_data_display)
+```
+
+## Health and personal care
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Healthcare Value and All-items Value Over Time"
+
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Health and personal care" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Health and personal care', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: tbl-1
+#| tbl-cap: "Comparison of Healthcare Value and All-items Value Over Time"
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Health and personal care', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, 'Health and personal care', `All-items`) |> 
+  rename(
+    `Healthcare Value` = 'Health and personal care', 
+    `All-items Value` = 'All-items'
+  )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+kable(wide_data_display)
+```
+## Recreation, education and reading
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-1
+#| fig-cap: "Comparison of Recreation, education and reading Value and All-items Value Over Time"
+
+# Load necessary libraries
+library(ggplot2)
+library(readr)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Filter for the "Recreation, education and reading" and "All-items" product groups
+filtered_data <- data[data$products_and_product_groups %in% c('Recreation, education and reading', 'All-items'),]
+
+# Plotting
+ggplot(filtered_data, aes(x = ref_date, y = value, color = products_and_product_groups, group = products_and_product_groups)) +
+  geom_line(size = 1) +  # Draw lines for both groups
+  geom_point(size = 2) +  # Highlight data points for both groups
+  labs(
+    y = "Value",
+    x = "Year",
+    color = "Product Group"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = filtered_data$ref_date) # Ensure all years are shown
+
+```
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: tbl-1
+#| tbl-cap: "Comparison of Recreation, education and reading Value and All-items Value Over Time"
+# Read and prepare the dataset for displaying in a table format
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
+
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
+
+# Reshape data to wide format using modern pipe syntax
+wide_data <- data |> 
+  filter(products_and_product_groups %in% c('Recreation, education and reading', 'All-items')) |>
+  spread(key = products_and_product_groups, value = value) |>
+  rename(Year = ref_date)
+
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+
+# Prepare the data for the table display
+wide_data_display <- wide_data |>
+  select(Year, 'Recreation, education and reading', `All-items`) |> 
+  rename(
+    `Recreation, education and reading Value` = 'Recreation, education and reading', 
+    `All-items Value` = 'All-items'
+  )
+
+# Render the table using knitr::kable and kableExtra::kable_styling
+kable(wide_data_display)
 ```

### 2024-04-18 scripts/99-replications.R

@@ -1,19 +1,25 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+library(readr)
+library(dplyr)
+library(knitr)
+library(kableExtra)
 
+# Read the dataset
+data <- read.csv(here::here("data/cleaned_data.csv"))
 
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
+# Reshape data to wide format
+wide_data <- data %>% 
+  filter(products_and_product_groups %in% c('Food', 'All-items')) %>%
+  spread(key = products_and_product_groups, value = value) %>%
+  rename(Year = ref_date)
 
-#### Load data ####
-# [...ADD CODE HERE...]
+# Assuming that the 'value' columns are already in the correct format and do not need additional formatting
+# If formatting is needed, include the appropriate mutate() calls here
 
+# Create the table
+wide_data_display <- wide_data %>%
+  select(Year, `Food`, `All-items`) %>% 
+  rename(`Food Value` = Food, `All-items Value` = `All-items`)
 
-
+# Render the table using knitr::kable and kableExtra::kable_styling
+wide_data_display %>%
+  kable()

### 2024-04-18 scripts/99-replications.R
test

@@ -23,3 +23,36 @@ wide_data_display <- wide_data %>%
 # Render the table using knitr::kable and kableExtra::kable_styling
 wide_data_display %>%
   kable()
+
+
+```{r}
+#| messages: false
+#| echo: false
+#| warning: false
+#| label: fig-2
+#| fig-cap: "Cost of Bacon Rate vs. Inflation Rate, data (since 2018)"
+#| fig-align: center
+#| fig-width: 10
+#| fig-height: 3
+#### Create graphs plotting mandated and actual response times ####
+bacon_table_display <- bacon_inflation_data |>
+  # Order the columns 
+  select(year, bacon, percent_change, inflation_rate) |> 
+  
+  # Change months to fiscal quarters
+  mutate(bacon =  paste("\\$", bacon, sep = "")) |>
+  mutate(percent_change =  paste(percent_change, "\\%", sep = "")) |>
+  mutate(inflation_rate =  paste(inflation_rate, "\\%", sep = "")) |>
+  
+  # Rename the columns
+  rename(
+    "Year" = "year", 
+    "Bacon Cost" = "bacon", 
+    "Inflation Rate" = "inflation_rate", 
+    "Bacon Cost Rate" = "percent_change"
+  )
+
+bacon_table_display |>
+  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "" )|>
+  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE)
+```

### 2024-04-20 outputs/paper.qmd
test

@@
 number-sections: TRUE
-bibliography: references.bib
+bibliography: outputs/references.bib
 thanks: "See the GitHub repository supporting this analysis: https://github.com/qinheinfo/Canada_inflation"
@@
 
+\newpage
+
 # Introduction
+The decision to analyze Canadian inflation during the COVID-19 pandemic was prompted by the significant economic disturbances observed globally.
+As a G7 nation with a prominent economic footprint, Canada offers a nuanced perspective on the complexities of inflation during times of crisis.
+The Consumer Price Index (CPI), a critical economic indicator, measures the average change over time in the prices paid by urban consumers for a standard basket of goods and services [@investopedia_CPI].
+The CPI is instrumental in shaping economic policies, such as adjustments to interest rates, wages, and social benefits, which in turn directly influence the everyday economic experiences of individuals [@Forum_CPI].
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Amidst the pandemic and subsequent global events like the war in Ukraine, consumer prices have been significantly impacted.
+Canadians have grappled with rising costs in many aspects of their daily lives due to supply chain disruptions, oil price volatility, and shifts in consumer demand.
+The CPI, Statistics Canada's primary measure of inflation, has captured these extraordinary impacts, with inflation rates reaching four-decade highs in 2022.
+[@Statisticsca_CPI]
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper delves into the relationship between the prices of various consumer goods and Canada's inflation rate.
+Leveraging data from Statistics Canada, it examines the five-year cost trajectory of common grocery store items across categories such as food, household goods, clothing, gasoline, healthcare, and recreation.
+Additionally, it analyzes yearly average inflation rates from 2019 to 2023.
+By studying the percentage increases in the prices of these items, this research seeks to uncover potential correlations between specific cost fluctuations and broader inflation trends, thereby enhancing our understanding of inflation's influence on the Canadian economy.[@CPI_data]
 
@@
 ## Food
+@fig-1 illustrates the trajectory of both the Food Value and the All-items Value over a five-year span, from 2019 to 2023.
+A stark contrast is evident in the divergence of these two lines post-2020, potentially indicative of the varying impact of the pandemic on food prices compared to general inflation.
+The steeper gradient of the Food curve suggests an amplified reaction to pandemic-related factors, which may include disruptions in the supply chain, changes in consumer demand, and perhaps shifts in agricultural productivity.
 
@@
 ```
+@tbl-1 offers a numerical perspective on the trends highlighted in Figure 1, providing the exact annual values that reflect the differential inflation rates for Food and All-items.
+It is clear from the table that the Food Value consistently increased at a higher rate than the All-items Value each consecutive year.
+Notably, between 2022 and 2023, the Food Value rose from 171.5 to 184.3, while the All-items Value exhibited a more moderate climb from 151.2 to 157.1.
+This quantifiable data underscores the disproportionate inflationary pressures on food items.
+
 ```{r}
@@
 ```
+both @fig-1 and @tbl-1 reveal that food-related inflation has outstripped the general inflation rate during the period analyzed.
+This divergence is most pronounced in the later years, notably between 2022 and 2023, suggesting that while the general cost of living increased during the pandemic, the cost of food increased even more markedly.
+These findings could have implications for economic policy, as they underscore the need to address the specific drivers of food inflation in the context of the pandemic's broader economic impact.
 
 ## Household operations, furnishings and equipment
+In @fig-2 depicts a comparative analysis specifically focusing on the Household Value, relative to the All-items Value, over a period from 2019 to 2023.
+The value associated with Household operations, furnishings, and equipment shows a gradual increase over the five-year span.
+The progression is steady from 2019 to 2021, followed by a more noticeable rise from 2021 to 2023.
+This suggests a delayed but eventual response to inflationary pressures within the household sector during and post the pandemic period.
+The consistency in the upward trajectory, however, does not match the sharp incline observed in the All-items Value, especially from 2021 onwards.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-1
+#| label: fig-2
 #| fig-cap: "Comparison of Household Value and All-items Value Over Time"
@@
 
-# Filter for the "Food" and "All-items" product groups
+# Filter for the "Household" and "All-items" product groups
 filtered_data <- data[data$products_and_product_groups %in% c('Household operations, furnishings and equipment', 'All-items'),]
@@
 ```
+@tbl-2 supports the observations from the graph by providing the exact figures for the Household Value.
+With a modest rise from 123.6 in 2019 to 132.0 in 2023, the increase in household-related costs is apparent, yet not as drastic as the general inflation trend.
+The year-to-year change remains relatively ,moderate, with the most significant jump observed between 2021 and 2022, suggesting a particular period where inflation hit household items more significantly compared to the preceding years.
+
 ```{r}
@@
 #| warning: false
-#| label: tbl-1
+#| label: tbl-2
 #| tbl-cap: "Comparison of Household Value and All-items Value Over Time"
@@
 ```
+In light of the inflation dynamics in Canada during and after the COVID-19 pandemic, the Household Value data illustrates that this category has experienced inflation, but not to the extreme extents seen in the aggregate of all items.
+This could indicate that household operations, furnishings, and equipment were subjected to different economic forces or benefited from factors that mitigated more extreme price increases, such as alternative supply chains, less volatile demand, or different pricing strategies within the sector.
 
 ## Clothing and footwear
+In @fig-3, the line graph showcases the 'Clothing and footwear' value relative to the 'All-items' value over a five-year period, from 2019 to 2023.
+The 'Clothing and footwear' line remains remarkably flat, indicating that the value has not significantly increased over time.
+This stability suggests that clothing and footwear were less affected by inflationary pressures compared to the average of all consumer goods and services, which is depicted by the rising 'All-items' line.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-1
+#| label: fig-3
 #| fig-cap: "Comparison of Clothing Value and All-items Value Over Time"
@@
 ```
+Delving into the numerical details provided in @tbl-3, we see that the Clothing Value begins at 96.1 in 2019 and shows a slight decrease until 2021, where it reaches its lowest at 94.1.
+This could indicate a reduced demand for clothing and footwear during the peak periods of the COVID-19 pandemic, possibly due to lockdowns and shifts in consumer spending towards essentials.
+However, by 2023, the Clothing Value modestly increases to 96.4, suggesting a slow return towards pre-pandemic levels.
+
 ```{r}
@@
 #| warning: false
-#| label: tbl-1
+#| label: tbl-3
 #| tbl-cap: "Comparison of Clothing Value and All-items Value Over Time"
@@
 ```
+The juxtaposition of the flat trend for Clothing and Footwear against the rising trend of the All-items Value highlights a distinct economic behavior within the clothing sector.
+This sector appears to have been shielded from the inflation that affected other sectors more broadly.
+Possible reasons for this could include a decrease in consumer spending on fashion and discretionary items during the pandemic, shifts in retail and e-commerce strategies, or an oversupply in the market.
 
 ## Gasoline
+@fig-4 presents a line graph comparing the Gasoline Value to the All-items Value from 2019 to 2023.
+The Gasoline line shows a significant and volatile shift over the period, with a notable dip in 2020 followed by a sharp increase through to 2022.
+This volatility is contrasted against the more gradual upward trend of the All-items Value.
+
+The steep decrease in Gasoline Value in 2020 aligns with the global reduction in oil prices and travel activity during the early stages of the COVID-19 pandemic.
+However, the subsequent sharp rise through 2021 and into 2022 might reflect the resurgence in travel demand, combined with fluctuations in oil production and supply chain issues, leading to a spike in gasoline prices that outpaces the general inflation rate.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-1
+#| label: fig-4
 #| fig-cap: "Comparison of Gasoline Value and All-items Value Over Time"
@@
 ```
+@tbl-4 supports the graphical data by providing the exact figures for both Gasoline and All-items Values.
+Starting at 170.5 in 2019, the Gasoline Value decreases to 146.7 in 2020.
+The year 2021 marks the beginning of a steep climb to 192.4, continuing to a peak at 247.2 in 2022 before a slight retreat to 228.4 in 2023.
+The All-items Value, in comparison, rises steadily from 136.0 in 2019 to 157.1 in 2023, without the same level of fluctuation seen in gasoline prices.
+
 ```{r}
@@
 #| warning: false
-#| label: tbl-1
+#| label: tbl-4
 #| tbl-cap: "Comparison of Gasoline Value and All-items Value Over Time"
@@
 ```
+Gasoline prices experienced significant fluctuations, with a dramatic increase that far exceeds the inflation trend of other consumer goods and services.
+The sharp rise in gasoline prices could have a cascading effect on the broader economy due to the essential nature of fuel in transportation and logistics, potentially contributing to the inflation of other items.
+The trend in gasoline prices is indicative of the unique economic pressures faced in the energy sector, including the impact of global events on oil markets and the transition to alternative energy sources.
 
 ## Health and personal care
+@fig-5 displays the value of 'Health and personal care' relative to the 'All-items' value over the span of 2019 to 2023.
+The line for 'Health and personal care' exhibits a steady upward trend, though at a slower rate than the 'All-items' value.
+The growth in the health sector is consistent, which could suggest ongoing inflationary pressures in health-related goods and services, but it is not as steep as the overall inflation.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-1
+#| label: fig-5
 #| fig-cap: "Comparison of Healthcare Value and All-items Value Over Time"
@@
 ```
+The numerical data in @tbl-5 offers a clearer picture of this trend.
+Starting at a value of 127.4 in 2019, 'Health and personal care' experiences a moderate increase each year, reaching 145.8 in 2023.
+The annual growth is steady without any significant spikes, indicating that while there has been inflation within the healthcare sector, it has been relatively controlled and predictable.
+
 ```{r}
@@
 #| warning: false
-#| label: tbl-1
+#| label: tbl-5
 #| tbl-cap: "Comparison of Healthcare Value and All-items Value Over Time"
@@
 ```
+This data may point to the fact that despite the overall economic turbulence experienced during the COVID-19 pandemic, the health and personal care sector has not inflated to the same extent as the aggregate of all items.
+It may also reflect the essential status of healthcare, which could lead to more stable pricing regardless of broader economic conditions.
+The relative stability in health and personal care costs could be critical for policy discussions around healthcare affordability and accessibility during times of economic stress.
+
 ## Recreation, education and reading
+In @fig-6, the trend for 'Recreation, education, and reading' demonstrates a gradual yet consistent upward movement over the years 2019 to 2023.
+It shows a steady increase in value, albeit not as steep as the 'All-items' category, which experiences a more pronounced rise during the same period.
+This suggests that while general inflation has been rising, the costs within the recreation and education sectors have been increasing at a comparatively milder rate.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-1
+#| label: fig-6
 #| fig-cap: "Comparison of Recreation, education and reading Value and All-items Value Over Time"
@@
 ```
+
+@tbl-6 offers a year-by-year comparison of the values for 'Recreation, education, and reading' against the 'All-items' values.
+Beginning in 2019, the value for recreation and education starts at 116.8, dips slightly to 115.7 in 2020, and then gradually increases to 117.9 in 2021.
+This slight decrease followed by an increase could be indicative of the initial impact of the pandemic, with subsequent adaptation and recovery in the sector.
+Moving into 2022 and 2023, the values continue to rise to 123.5 and 126.4, respectively, underscoring a continued but moderate growth in contrast to the more significant inflationary trend seen across the broad spectrum of consumer goods and services, where the 'All-items' value moves from 136.0 in 2019 to 157.1 in 2023.
+
 ```{r}
@@
 #| warning: false
-#| label: tbl-1
+#| label: tbl-6
 #| tbl-cap: "Comparison of Recreation, education and reading Value and All-items Value Over Time"
@@
 ```
+Recreation, education, and reading sector saw an overall increase in value throughout the five-year period, although at a pace that was slower than the overall increase in consumer prices.
+The resilience of this sector's value, rising despite the economic challenges of the pandemic, may suggest a robust demand for recreational, educational, and reading products and services.
+Additionally, it could reflect the impact of digital alternatives and online services that became more prevalent during the pandemic, potentially offsetting higher costs or changes in consumer behavior as individuals and families sought at-home entertainment and learning options.
 
 # Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 

### 2024-04-20 outputs/paper.qmd
Update

@@
 number-sections: TRUE
-bibliography: outputs/references.bib
+bibliography: references.bib
 thanks: "See the GitHub repository supporting this analysis: https://github.com/qinheinfo/Canada_inflation"
@@
 ## Measurement
+The data used in this paper comes from Statistics Canada’s Consumer Price Index (CPI), Prices for the CPI are collected across a variety of geographic locations and from a selection of goods and services representative of Canadian consumer spending.
+The data collection targets urban and rural private households, excluding specific demographics such as those residing in collective households or on Indian reserves.
+Price collectors employ a detailed specification approach to ensure consistency in the quality and type of goods and services priced [@CPI_info].
+
+Sampling for the CPI is carefully designed to reflect the relative importance of each item in the total expenditures of consumers.
+A mix of probability and non-probability sampling techniques is employed to select a diverse and representative set of retail outlets and goods.
+Prices are typically collected during the first two weeks of each reference month, with some variations depending on the nature of the goods or services.
+
+The CPI data are gathered from direct survey responses, administrative records, and other relevant Statistics Canada surveys.
+This multimodal collection ensures a comprehensive capture of price movements across the selected basket items.
+Prices are collected monthly, quarterly, semi-annually, or annually based on the nature of the goods or services, with food prices often collected into the third week of the month.
 
 ## Data Analysis
-This paper uses data from both Statistics Canada’s Consumer Price Index (CPI) and the Bank of Canada’s Inflation Rate measurements.
-Within the CPI, the paper specifically examines 6 variables representing different food items: bacon (per 500 grams), eggs (per 1 dozen), chicken breasts (per kilogram), milk (per 2 liters), frozen peas (per 750 grams), and pasta sauce (per 650 milliliters).
-These 6 food items were 3selected as a representation of various food types and groups, allowing an analysis of meats, dairy, frozen goods, and shelf-stable food.
-The CPI displays a price of each food item by month of a given year, and was subsequently cleaned to find the average of a given year.
-Limitations in the history of the data only allow for analysis from 2017 up to 2022.
-Once the data was cleaned, a cost rate (the difference in price of a given year against the previous year) was calculated to show the change in price in order to measure the food item against inflation.
-It is important to note that, since the data only extends to 2017, a cost rate analysis from the years 2018 to 2022.
-The rate of inflation measurements similarly recorded data month-to-month and required cleaning to show the yearly average.
-Although the inflation rate data set provided a far more robust history, only the years 2017 to 2022 were included in this analysis as inflation was only used as a comparison against food cost rates.
+This paper utilizes data from Statistics Canada's Consumer Price Index (CPI), focusing on a subset of six variables that represent diverse categories of consumer spending: food, household essentials, clothing, gasoline, healthcare, and recreation.
+These categories were carefully selected to encompass a broad spectrum of goods and services, providing a comprehensive sight of consumer price changes during significant economic periods.
+The analysis covers the period from 2019 to 2023, a timeline strategically chosen to encompass the main phases of the COVID-19 pandemic as well as the subsequent post-pandemic economic environment.
 
@@
 
+## Summary
+The overall CPI rose steadily from 2019 to 2023, peaking in 2022, indicative of broad economic pressures.
+This general trend highlights the pervasive impact of the pandemic across all sectors, though the intensity of the impact varied.
+The analysis underscores the complex interplay between supply chain disruptions, altered consumer behaviors, and inflationary expectations, which collectively influenced the trajectory of price changes.
+
+The varied responses across sectors suggest differential resilience and vulnerability to economic shocks.
+For example, the sharp rise in food prices not only reflects supply chain issues but also changes in consumer preferences and possible speculative behaviors.
+Similarly, the fluctuations in gasoline prices highlight the external dependency on global markets, which can rapidly transmit economic shocks into domestic inflation.
+
+## Limitations
+The primary constraint is the reliance on CPI data, which, despite its comprehensive nature, may not capture the granular impacts of economic policies, consumer sentiment, or unrecorded market activities.
+Moreover, the analysis period, though strategically chosen to cover the main phases of the pandemic, does not allow for the examination of long-term economic impacts, which may become more apparent in subsequent years [@IMF_blog].
+
+## Future Directions
+Future research should aim to address these limitations by incorporating more diverse data sources, such as consumer surveys or micro-level transaction data, to provide a deeper understanding of individual and sector-specific economic behaviors [@UTokyo].
+Longitudinal studies extending beyond the immediate aftermath of the pandemic could offer insights into the lasting effects of such a global crisis on consumer price indices [@IMF_blog].
+
+Additionally, comparative studies involving multiple countries could enhance the robustness of the findings, allowing for a broader evaluation of CPI dynamics in different economic systems and stages of recovery [@Global_inflation].
+Exploring the impact of specific government interventions or policy changes on the CPI could also yield valuable insights into effective strategies for managing inflation in crisis situations.
+
 \newpage

### 2024-04-20 outputs/paper.qmd
Update

@@
 bibliography: references.bib
-thanks: "See the GitHub repository supporting this analysis: https://github.com/qinheinfo/Canada_inflation"
+thanks: "The GitHub repository: https://github.com/qinheinfo/Canada_inflation"
 ---
@@
 
-This paper delves into the relationship between the prices of various consumer goods and Canada's inflation rate.
-Leveraging data from Statistics Canada, it examines the five-year cost trajectory of common grocery store items across categories such as food, household goods, clothing, gasoline, healthcare, and recreation.
-Additionally, it analyzes yearly average inflation rates from 2019 to 2023.
-By studying the percentage increases in the prices of these items, this research seeks to uncover potential correlations between specific cost fluctuations and broader inflation trends, thereby enhancing our understanding of inflation's influence on the Canadian economy.[@CPI_data]
+This paper explores the relationship between the CPI values of specific product categories and the all-items CPI.
+Utilizing data from Statistics Canada, it examines the cost trajectory from 2019 to 2023 for commonly purchased items in grocery stores across categories such as food, household goods, clothing, gasoline, healthcare, and recreation.
+By analyzing the increases in CPI values for these items, this research aims to identify potential correlations between specific cost fluctuations and broader inflation trends.
+This analysis enhances our understanding of how inflation impacts the Canadian economy, shedding light on the dynamic interplay between individual spending categories and overall economic conditions.
+[@CPI_data]
 

### 2024-04-21 outputs/paper.qmd
Update

@@
 
+## Dataset
+
+The Dataset were derived from Statistics Canada's Consumer Price Index (CPI), capturing key statistics on Canadian Consumaer Price Index Value of different items encompassing from 2019 to 2023.
+
+@tbl-7 shows the first 5 rows in the dataset, REF_DATE stands for the time that the data was gathered, products_and_product_groups stands for items category, value stands for CPI value.
+
+```{r}
+#| label: tbl-7
+#| tbl-cap: First 5 Rows in Mental Health Indicators Dataset 
+#| echo: false
+
+kable(head(product_data,5))
+```
+
 ## Data Analysis
+To analyze the data , I worked on R [@citeR], using the tidyverse [@tidyverse], [@arrow], and gitcreds [@gitcreds] packages.
+Tables in the report were created with kableExtra [@kableExtra] and figures were created with ggplot2 which was built in tidyverse [@tidyverse].
+
 This paper utilizes data from Statistics Canada's Consumer Price Index (CPI), focusing on a subset of six variables that represent diverse categories of consumer spending: food, household essentials, clothing, gasoline, healthcare, and recreation.
@@
 ```
-@tbl-1 offers a numerical perspective on the trends highlighted in Figure 1, providing the exact annual values that reflect the differential inflation rates for Food and All-items.
+@tbl-1 offers a numerical perspective on the trends highlighted in @fig-1, providing the exact annual values that reflect the differential inflation rates for Food and All-items.
 It is clear from the table that the Food Value consistently increased at a higher rate than the All-items Value each consecutive year.
